Predicting Sequence Variability from Voronoi Tessellation of Proteins

### 2015-03-27 manuscript/voronoi/mnv.tex
MN modified significantly, new figures in png format added

\documentclass[11pt]{article}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{nicefrac}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage{footnote}
\usepackage{rotating}
\usepackage{slashbox}
\usepackage{afterpage}
\usepackage{float}
\citestyle{mbe}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%force setting figures at the top of the page:
    \makeatletter
    \setlength{\@fptop}{0pt}
    \makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  personal abbreviations and macros
%    the following package contains macros used in this document:
%\input econfmacros.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textwidth=6.9in  \textheight=9.6in
%%  Adjust these for your printer:
\leftmargin=-5in   \topmargin=-0.8in
\hoffset = -0.95in  \voffset = 0pt
\marginparwidth = 0pt
%% preprint number data:
%\newcommand\pubnumber{Article 14 in eConf C1304143}
\newcommand\pubnumber{}
\newcommand\pubdate{\today}

@@

    Motivated by the existing gaps in the current understanding of the role of flexibility and other structural properties on sequence-structure relations in proteins, here we propose and derive a new set of site-specific structural properties which, unlike CN and WCN, their definitions does not involve any free parameters, while performing equally well or better than all previously-considered structural quantities in predicting protein sequence evolution. This is done by employing tessellation methods from the field of computational geometry to calculate several new characteristics of sites in proteins, which can serve as proxy measures of local packing density and site-specific flexibility. Contrary to what is currently perceived about the role of flexibility in sequence variability \citep[e.g.,][]{huang_mechanistic_2014}, we show that the newly calculated flexibility measures outperform many of previously studied structural properties, such as RSA and the traditional definitions of Contact Number and the Weighted Contact Number (WCN), in predicting sequence evolution at residue level. \\

@@

        Depending on the choice of coordinates used, there exist degeneracy in the definition of some site-specific structural variables. For example, the quantity WCN is generally calculated from the coordinates of $C_\alpha$ atoms in the 3-dimensional structure of protein. The choice of $C_\alpha$ coordinates is however mainly driven by convenience in WCN calculation and there is no reason to believe this set of atomic coordinates are the best representatives for individual sites in proteins. Indeed, some earlier works have already suggested the use of center-of-mass of side chain coordinates to represent the 3D structure of protein \citep[e.g., ][]{soyer_voronoi_2000}. More recently, \citet{marcos_too_2014} have also shown that WCN calculated from side-chain center-of-mass coordinates generally result in significantly better correlations of WCN with sequence variability measures.\\

@@

        In order to identify which set of atomic coordinates best represents individual sites for the calculation of WCN, B factor, and Voronoi cells, here we consider all possible choices of the representative set of atomic coordinates. These include the set of coordinates of all backbone atoms ($N$, $C$, $C_\alpha$, $O$) and the first heavy atom in the amino acid side chains ($C_\beta$). In addition, we calculate representative coordinates for each site in protein by averaging over the coordinates of all heavy atoms in the side chains. We also calculate a representative coordinate for each site by averaging over all heavy atom coordinates in the side chain and the backbone of the amino acid together. In rare cases where the side chain $C_\beta$ atom had not been resolved in the PDB file or the amino acid lacked $C_\beta$ (e.g., Glycine), the $C_\beta$ coordinate for the specific amino acid were replaced with the coordinate of the corresponding $C_\alpha$ atom in the same amino acid. \\

@@

    \subsection*{Voronoi Cell Area and Volume as a Measure of Local Packing Density in Proteins}

    In order to assess the prediction power site-specific variables derived from Voronoi tessellation, we first calculate the geometric centers of all side-chains for each of the proteins in the dataset and use them as the seeds of Voronoi polyhedra. Plot {\it A} of Figure \ref{fig:best_voronoi} depicts the distributions of the Spearman's correlation coefficients of five most important Voronoi cell characteristics with site-specific evolutionary rates (ER). It should be noted that all cell characteristics in the plot correlate positively with ER, except the cell sphericity which is always negatively correlated with ER and other Voronoi cell properties. In general, we observe that the cell surface area has the best prediction power compared to other cell characteristics, followed by the cell volume, cell eccentricity as defined in previous section, cell's total edge length, and the cell sphericitiy. The cell properties are also highly associated with each other. Although the Voronoi cell volume is the second best correlating variable with ER, it exhibits no significant independent correlation with ER once we control for the cell area with the median of the distribution centered at $\sim0.0$, as illustrated in plot {\it B} of Figure \ref{fig:best_voronoi}. On the contrary, the cell sphericity and eccentricity both exhibit median partial correlations of $\sim-0.1$ \& $\sim0.07$ with ER respectively, when the contribution from the Voronoi cell area is controlled. In conclusion, the cell area, volume, and edge length appear to almost represent the same property of the Voronoi cell. Other Voronoi cell characteristics, such as the number of vertices, faces and edges of the cell also tend to correlate weakly with sequence evolutionary rates. These cell characteristics are however, discrete (integer) quantities and in general have a limited range.

    Not shown here for brevity, we also obtain identical results to the above if we use sequence entropy as defined by Eqn. \ref{eqn:shannon} instead of sequence evolutionary rates, although sequence entropy generally results in weaker correlation strengths due to the discreteness and limited range inherent in its definition.

@@

        Depending on the set of atomic coordinates that represent the protein structure, there are $7$ different measures for site-specific structural properties such as Weighted Contact Number, B factor and Voronoi cell properties. Therefore, in order to eliminate redundant variables from dataset, we first compare the predictive power of different measures of residue characteristics based on the set of atomic coordinates used.
        \\

        For the measure of local packing density in proteins (the Weighted Contact Number) we find that among all possible set of coordinates, the average over coordinates of all heavy atoms of each individual side chain results in WCN values that show the best correlation with other structural and sequence properties, such as RSA, Voronoi cell properties, sequence entropy, and evolutionary rates. Specifically, WCN from average side chain coordinates (wcnSC) outperforms WCN based on $C_\alpha$ coordinates (wcnCA) in predicting RSA, \ddg entropy, sequence entropy and evolutionary rates (r4sJC) by a median Spearman correlation difference of $0.09$, $0.10$, $0.07$ \& $0.08$, respectively (Figure \ref{fig:best_wcn}).
        \\

        For the measure of local flexibility in proteins (B factor) we similarly find that among all $7$ representative measures of site B factors, the average of B factor values over all heavy atoms of each individual side chain (bfSC) results in the best correlations with other structural and sequence properties. Specifically, bfSC outperforms the commonly used $C_\alpha$ B factor (bfCA) in predicting RSA, \ddg entropy, sequence entropy and evolutionary rates by a median Spearman correlation difference of $0.11$, $0.12$, $0.08$ \& $0.09$, respectively (Figure \ref{fig:best_bf}).
        \\

@@

\section{Discussion and Concluding Remarks}
\label{sec:dcr}

		Throughout the previous sections, a new set of parameter-free site-specific structural quantities were introduced using Voronoi tessellation that are capable of explaining sequence evolutionary rates equally or better than many of the previously considered site-specific structural characteristics, including RSA, ddG rate as defined by \citet{echave_relationship_2014}, site-specific measures of flexibility such as B factor, and the traditional definition of Contact Number and the Weighted Contact Number using $C_\alpha$ atomic coordinates. It is however notable that, once we recalculate WCN using the geometric center of the side chains as the representative coordinates of individual sites, the quantity WCN still outperforms all other structural quantities, including those derived from Voronoi tessellation, in explaining site-specific evolutionary rates.

        All observations clearly demonstrate that individual sites in proteins are best represented by the average properties of the side chains of amino acids in the corresponding sites. In particular, the strength of structure and sequence correlations decrease when moving from side chain to backbone atoms. An exception to this general pattern is the correlation of the hydrogen-bond energies of the sites, which correlate more strongly with site characteristics calculated based on the backbone atoms instead of side chain.
        \\

        Based on the observations described in the previous paragraphs, we keep only variables measured from average side chain properties and coordinates throughout the rest of the analysis and omit all other similar measures that show only weaker correlations with other site-specific characteristics. The exclusion of these alternative measures results in a significant reduction in the number of pdb-level variables to be further analysed, without compromising generality and comprehensiveness of the analysis.
        \\

		One potential caveat with Voronoi tessellation is the edge effects, that is, Voronoi cells that are on the surface of the protein are open. To ensure that

        Throughout this work, we have carried out a comprehensive analysis in search for the main determinants of the strength of sequence--structure correlations -- some of which are newly reported and discussed in this work. Examples of sequence--structure relations include the correlations of sequence entropy ({\it seqent}) and measures of evolutionary rates (such as {\it r4sJC} used in this work) with measures of residue Contact Number (e.g., {\it wcnSC}), Relative Solvent Accessibility (RSA), \ddg entropy ({\it ddgent}). In addition, we have derived new site-specific properties, based on Voronoi Tessellation of protein 3D structures, that are comparable to or better than several previously known structural properties in explaining site-specific sequence entropy or evolutionary rates (e.g., Figure \ref{fig:seqent_structure_cors}). Prime examples include Voronoi cell volume ({\it vvolume}), surface area ({\it varea}) and

         We have also shown that site-specific structural properties -- such as Weighted Contact Number, Bfactor and Voronoi Cell properties -- that are calculated from the average coordinates of side chain atoms, have the best explanatory powers for the sequence variability measures such as {\it seqent} and {\it r4sJC}. Compared to the common choice of back bone {\it CA} atomic coordinates, site-specific properties averaged over side chain atoms can outperform in predicting sequence evolutionary rates by as much as $0.12$ in terms of Spearman correlation strength.

        In search for the determinants of the strength of sequence-structure relations, we compiled a set of more than $200$ protein properties for a dataset of $209$ monomeric enzymes. By employing several independent parametric and non-parametric statistics, such as Spearman rank test, regularized regression and Principal Regression methods, we identify sequence divergence as the dominant factor in the strength of sequence-structure correlations, capable of explaining $10-30\%$ of the observed correlation strengths alone, in both the original and rank-transformed data.

@@

We thank Austin G. Meyer, Stephanie Spielman and Eleisha Jackson at UT Austin for helpful discussions.

@@

>>>>

\documentclass[11pt]{article}
\usepackage{amssymb}
\usepackage{hyperref}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{nicefrac}
\usepackage[usenames,dvipsnames]{xcolor}
\usepackage{graphicx}
\usepackage{footnote}
\usepackage{rotating}
\usepackage{slashbox}
\usepackage{afterpage}
\usepackage{float}
\citestyle{mbe}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%force setting figures at the top of the page:
    \makeatletter
    \setlength{\@fptop}{0pt}
    \makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%  personal abbreviations and macros
%    the following package contains macros used in this document:
%\input econfmacros.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\textwidth=6.9in  \textheight=9.6in
%%  Adjust these for your printer:
\leftmargin=-5in   \topmargin=-0.8in
\hoffset = -0.95in  \voffset = 0pt
\marginparwidth = 0pt
%% preprint number data:
%\newcommand\pubnumber{Article 14 in eConf C1304143}
\newcommand\pubnumber{}
\newcommand\pubdate{\today}

@@

    Motivated by the existing gaps in the current understanding of the role of flexibility and other structural properties on sequence-structure relations in proteins, here we propose and derive a new set of site-specific structural properties which, unlike CN and WCN, their definitions does not involve any free parameters, while performing equally well or better than all previously-considered structural quantities in predicting protein sequence evolution. This is done by employing tessellation methods from the field of computational geometry to calculate several new characteristics of sites in proteins, which can serve as proxy measures of local packing density and site-specific flexibility. Contrary to what is currently perceived about the role of flexibility in sequence variability \citep[e.g.,][]{huang_mechanistic_2014}, we show that the newly calculated flexibility measures outperform many of previously studied structural properties, such as RSA and the traditional definitions of Contact Number (CN) and the Weighted Contact Number (WCN), in predicting sequence evolution at residue level. \\

@@

        Depending on the choice of coordinates used, there exist degeneracy in the definition of some site-specific structural variables. For example, the quantity WCN is generally calculated from the coordinates of $C_\alpha$ atoms in the 3-dimensional structure of protein. The choice of $C_\alpha$ coordinates is however mainly driven by convenience in WCN calculation and there is no reason to believe this set of atomic coordinates is the best representative of individual sites in proteins. Indeed, some earlier works have already suggested the use of center-of-mass of side chain coordinates to represent the 3D structure of protein \citep[e.g., ][]{soyer_voronoi_2000}. More recently, \citet{marcos_too_2014} have also shown that WCN calculated from side-chain center-of-mass coordinates generally result in significantly better correlations of WCN with sequence variability measures.\\

@@

         Here we consider all possible choices of the representative set of atomic coordinates in order to identify which set of atomic coordinates best represents individual sites for the calculation of WCN, B factor, and Voronoi cells. These include the set of coordinates of all backbone atoms ($N$, $C$, $C_\alpha$, $O$) and the first heavy atom in the amino acid side chains ($C_\beta$). In addition, we calculate representative coordinates for each site in protein by averaging over the coordinates of all heavy atoms in the side chains. We also calculate a representative coordinate for each site by averaging over all heavy atom coordinates in the side chain and the backbone of the amino acid together. In rare cases where the side chain $C_\beta$ atom had not been resolved in the PDB file or the amino acid lacked $C_\beta$ (e.g., Glycine), the $C_\beta$ coordinate for the specific amino acid were replaced with the coordinate of the corresponding $C_\alpha$ atom in the same amino acid. \\

@@

    \subsection*{Voronoi Cell Area and Volume as Proxy Measures of Local Packing Density and Flexibility in Proteins}

    In order to assess the prediction power of site-specific variables derived from Voronoi tessellation, we first calculate the geometric centers of all side-chains for each of the proteins in dataset and use them as the seeds of Voronoi polyhedra. Plot {\it A} of Figure \ref{fig:best_voronoi} depicts the distributions of the Spearman's correlation coefficients of five most important Voronoi cell characteristics with site-specific evolutionary rates (ER). It is notable that all cell characteristics in the plot correlate positively with ER, except the cell sphericity which is always negatively correlated with ER and other Voronoi cell properties. In general, we observe that the cell surface area has the best prediction power compared to other cell characteristics, followed by the cell volume, cell eccentricity as defined in previous section, cell's total edge length, and the cell sphericitiy. The cell properties are also highly associated with each other. Although the Voronoi cell volume is the second best correlating variable with ER, it exhibits no significant independent correlation with ER once we control for the cell area, with the median of its distribution centered at $\sim0.0$, as illustrated in plot {\it B} of Figure \ref{fig:best_voronoi}. On the contrary, the cell sphericity and eccentricity both exhibit median partial correlations of $\sim-0.1$ \& $\sim0.07$ with ER respectively, when the contribution from the Voronoi cell area is controlled. In conclusion, the cell area, volume, and edge length appear to almost represent the same property of the Voronoi cell. Other Voronoi cell characteristics, such as the number of vertices, faces and edges of the cell also tend to correlate weakly with sequence evolutionary rates. These cell characteristics are however, discrete (integer) quantities and in general have a limited range.

    Not shown here for brevity, we also obtain identical results to above if we use sequence entropy as defined by Eqn. \ref{eqn:shannon} instead of sequence evolutionary rates. The use of sequence entropy however, generally results in weaker correlation strengths due to the discreteness and limited range inherent in its definition.

@@

        Depending on the set of atomic coordinates that represent the protein structure, there are at least $7$ different measures of each individual  site-specific structural properties such as Weighted Contact Number, B factor and Voronoi cell properties. Therefore, in order to eliminate redundant variables from dataset, we first compare the predictive power of different measures of residue characteristics based on the set of atomic coordinates used.
        \\

        For the measure of local packing density in proteins (the Weighted Contact Number) we find that among all possible set of coordinates, the average over coordinates of all heavy atoms of each individual side chain results in WCN values that show the best correlation with other structural and sequence properties, such as RSA, Voronoi cell properties, sequence entropy, and evolutionary rates. Specifically, WCN from average side chain coordinates outperforms WCN based on $C_\alpha$ coordinates in predicting RSA, \ddg entropy, sequence entropy and evolutionary rates with median Spearman correlation differences of $0.09$, $0.10$, $0.07$ \& $0.08$, respectively (Figure \ref{fig:best_wcn}).
        \\

        For the measure of local flexibility in proteins (B factor) we similarly find that among all $7$ representative measures of site B factors, the average of B factor values over all heavy atoms of each individual side chain results in the best correlations with other structural and sequence properties. Specifically, the average side chain B factor outperforms the commonly used $C_\alpha$ B factor in predicting RSA, \ddg entropy, sequence entropy and evolutionary rates by a median Spearman correlation difference of $0.11$, $0.12$, $0.08$ \& $0.09$, respectively (Figure \ref{fig:best_bf}).
        \\

@@

\section{Discussion}
\label{sec:dcr}

        Throughout this work, we have carried out a comprehensive analysis and comparison of the main structural determinants of sequence variability, some of which are newly reported and discussed in this work. Examples of sequence--structure relations include the correlations of measures of evolutionary rates such as {\it r4sJC} used in this work, and sequence entropy with measures of residue Contact Number, Relative Solvent Accessibility (RSA), and \ddg Rate as defined by \citet{echave_relationship_2014} which is essentially a proxy measure of stability of protein's native conformation upon substitution of amino acids in individual sites in proteins. In addition, we have derived new site-specific characteristics from the Voronoi Tessellation of protein 3D structures, that are capable of explaining sequence variability equally well or better than several previously considered structural quantities, such as B factor, RSA, \ddg Rate, and the traditional definitions of Contact Number (CN) and the Weighted Contact Number (WCN) using $C_\alpha$ atomic coordinates (e.g., Figure \ref{fig:best_predictor}). \\

        One potential caveat with Voronoi tessellation of finite structures in Euclidean space is the {\it edge effects}. Sites that are close to the surface of protein are often associated with Voronoi cells that are bounded by the cubic box containing the protein. To ensure that these edge effects do not influence the observed sequence-structure correlations, we identified open cells -- i.e., cells that are partially bounded and closed by the cubic box containing the protein -- in all proteins by examining the variations in individual cell volumes upon changing the size of the cubic box containing the protein. The open cells in individual proteins were then ranked by the fraction of volume changes observed upon changing the box size and then normalized to the the largest volume observed among closed cells. \\

        Our conclusion is that {\it edge effects} due to Voronoi tessellation have $\lesssim0.01$ influence on the observed sequence-structure correlation strengths in the dataset of $209$ proteins considered in this work. Similar conclusions are reached if the open cells were alternatively ranked by the fractional changes in cell area (vs. cell volume) upon changing the box size. Our conclusion is that Voronoi cell characteristics, in particular cell volume and area can be safely used in predicting sequence variability without recourse to corrections for edge effects. An exception is cell sphericity as defined in Eqn. \ref{eqn:sphericity}, which turns out to behave differently for open and closed cells. This is well illustrated in the adjacent averaging plots of Figures \ref{fig:adjavg_ER} \& \ref{fig:adjavg_SE} where we plot the behavior open and closed cell characteristics averaged over all sites in all proteins in our dataset, versus sequence evolutionary rate and sequence entropy. \\

        The definitions and values of some site-specific structural quantities, such as WCN and Voronoi cell characteristics, depend on the set of atomic coordinates that are used for representing individual sites in proteins. In order to identify which set of atomic coordinates best represent protein 3-dimensional structure for the calculation of structural quantities, we calculated WCN and Voronoi cell properties using different sets of atomic coordinates representing individual sites in proteins. Moreover, there is also ambiguity as to which set of atomic B factors best represent the flexibility of individual sites in protein. \\

        Our finding is that sequence variability measures correlate best with site-specific structural quantities such as WCN and Voronoi cell characteristics if the geometric center of the side-chains of amino acids are used for the calculation of these structural quantities instead of using the more commonly adopted backbone $C_\alpha$ atomic coordinates. In addition, the average of side-chain atomic B factors appear to correlate significantly better with sequence variability measures and site-specific structural quantities such as Relative Solvent Accessibility, \ddg Rate, and measures of amino acid hydrophobicities as illustrated in the plots of Figure \ref{fig:best_bf}.

        All observations clearly demonstrate that individual sites in proteins are best represented by the average properties of the side chains of amino acids in the corresponding sites rather than by the properties and the spatial coordinates of the backbone $C_\alpha$ atoms. In particular, the strength of structure-structure and sequence-structure correlations decrease monotonically when moving from side chain to backbone atoms. An exception to this general pattern is the correlation of the hydrogen-bond energies of individual sites with other site-specific structural properties. In general, average site-specific H-bond energies correlate more strongly with representative B factors, contact number, and Voronoi cell characteristics, if these quantities are calculated using the backbone Oxygen atoms as representation of individual sites, instead of using the average side chain properties and coordinates. This is well illustrated in the bottom-right plots of Figures \ref{fig:best_voronoi}, \ref{fig:best_wcn}, \& \ref{fig:best_bf}. The observed monotonic increase in the correlation strengths with H-bond energies from side chain to backbone $O$ atom can be explained away knowing that the backbone Oxygen atom is responsible for virtually all Hydrogen bonds in proteins. The influence of individual atoms on H-bond energies monotonically decreases from the neighbor backbone atoms $C$ \& $N$ to the farthest atoms from Oxygen in the amino acid side chains. \\

        Regardless of the set of atomic coordinates used for representing individual sites in proteins and for calculating site-specific structural quantities it is notable that once WCN is recalculated using the same set of coordinates used for Voronoi tessellation of protein structure, the quantity WCN as defined by Eqn \ref{eqn:wcn_pwrl} outperforms all other structural quantities --  including those derived from Voronoi tessellation -- in explaining site-specific sequence variability measured, as depicted in Figure \ref{fig:best_predictor}. This might not be surprising, knowing that the mathematical definition of WCN includes all sites in proteins which are all likely involved in long-range correlations and electrostatic interactions. In contrast to WCN, measures of site flexibility, in particular, the Voronoi cell volume or its inverse as a more local measure of packing density than WCN, do not include the possible effects of long-range interactions among non-neighbor amino acids in proteins. In this regard, it is reasonable to expect the Voronoi cell area to perform slightly better than cell volume in explaining the general patterns of sequence evolutions, since the Voronoi cell area is primarily determined by the local packing in addition to the number of neighboring amino acids around a given site of interest in protein, whereas the cell volume is more representative of the local packing and site flexibility than the number of neighboring amino acids around the site of interest.

@@

We thank Austin G. Meyer, Stephanie Spielman and Eleisha Jackson at UT Austin for helpful discussions and comments.

@@

    \begin{figure}[tbh]
        \begin{center}
        \includegraphics[width=6.9in]{adjacent_averaging_screen/voro_log/zr4s_JC.png}
        \end{center}
        \caption{General behavior of site-specific structural characteristics versus site-specific evolutionary rates among all sites in all 209 proteins in dataset. For plots depicting evolutionary rates versus versus Voronoi cell characteristics, the red \& blue curves represent respectively the general behaviors of closed \& blue open Voronoi cells, in addition to the black curves which represent the behavior of all open and closed cells together. The amino acid hydrophobicity scales were taken from \citet{hessa_recognition_2005}.}
        \label{fig:adjavg_ER}
    \end{figure}

    \begin{figure}[tbh]
        \begin{center}
        \includegraphics[width=6.9in]{adjacent_averaging_screen/voro_log/seqent.png}
        \end{center}
        \caption{General behavior of site-specific structural characteristics versus site-specific sequence variability, as measured by sequence entropy (Eqn. \ref{eqn:shannon}), among all sites in all 209 proteins in dataset. For plots depicting sequence entropy versus Voronoi cell characteristics, the red \& blue curves represent respectively the general behaviors of closed \& blue open Voronoi cells, in addition to the black curves which represent the behavior of all open and closed cells together. The amino acid hydrophobicity scales were taken from \citet{hessa_recognition_2005}.}
        \label{fig:adjavg_SE}
    \end{figure}
